---
title: "hw1.rmd"
author: "Student ID:110071010"
output:
  word_document: default
  pdf_document: default
---

## Read customers.txt

```{r}
customers <- read.table("customers.txt", header = TRUE)
ages <- customers$age
```

## 1. What is the 5th element in the original list of ages?

```{r}
# Find the fifth element of the ages vector after reading in the table
ages[5]
```

## 2. What is the fifth lowest age?

```{r}
# Get the ages vector sorted in ascending order and find its fifth element
sort(ages)[5]
```

## 3. Extract the five lowest ages together

```{r}
# Get the ages vector sorted in ascending order and employ slicing to get the first five elements
sort(ages)[1:5]
```

## 4. Get the five highest ages by first sorting them in decreasing order first

```{r}
# Get the ages vector sorted in decreasing order first and employ slicing to get the first five elements
sort(ages, decreasing = TRUE)[1:5]
```

## 5. What is the average (mean) age?

```{r}
# Use mean() to compute mean value
mean(ages)
```

## 6. What is the standard deviation of ages?

```{r}
# Use sd() to compute standard deviation
sd(ages)
```

## 7. Make a new variable called age_diff, with the difference between each age and the mean age

```{r}
# Store the difference in age_diff. Then use head() to show part of the resulting vector
age_diff <- ages - mean(ages)
head(age_diff)
```

## 8. What is the average "difference between each age and the mean age"?

```{r}
# Use mean() to figure out the average difference, but it does not yield a zero
mean(age_diff)
```

## 9. Visualize the raw data as we did in class: (a) histogram (b) density plot (c) boxplot + stripchart

```{r}
# Use methods introduced in class to plot the required graphs
# (a) histogram
hist(ages)
# (b) density plot
plot(density(ages))
# (c) boxplot + stripchart
boxplot(ages, horizontal = TRUE)
stripchart(ages, method = "stack", add = TRUE)
```
