---
title: "bacs_hw8"
author: '110071010'
date: "2024-04-13"
output: word_document
---

# Question 1

We will use the interactive_regression() function from CompStatsLib again.

[**Question**]{.underline}

Comparing scenarios 1 and 2, which do we expect to have a stronger R\^2 ?

[**Answer**]{.underline}

Scenario 1

## 1b

[**Question**]{.underline}

Comparing scenarios 3 and 4, which do we expect to have a stronger R\^2 ?

[**Answer**]{.underline}

Scenario 3

## 1c

[**Question**]{.underline}

Comparing scenarios 1 and 2, which do we expect has bigger/smaller SSE, SSR, and SST? (intuitively)

[**Answer**]{.underline}

SSE : Scenario 2 has larger SSE than Scenario 1 does

SSR : Scenario 1 has larger SSE than Scenario 2 does

SST : likely to be slightly higher in Scenario 2 due to greater overall variance

## 1d

[**Question**]{.underline}

Comparing scenarios 3 and 4, which do we expect has bigger/smaller SSE, SSR, and SST? (intuitively)

[**Answer**]{.underline}

SSE : Scenario 4 has larger SSE than Scenario 3 does

SSR : Scenario 3 has larger SSE than Scenario 4 does

SST : likely to be slightly higher in Scenario 4 due to greater variance

# Question 2

## 2a

Use the lm() function to estimate the regression model Salary \~ Experience + Score + Degree. Show the beta coefficients, R\^2, and the first 5 values of y (\$fitted.values) and (\$residuals)

```{r, results="hold"}
salary_data <- read.table("programmer_salaries.txt", header=TRUE, sep="\t")
regr <- lm(Salary ~ Experience + Score + Degree, data=salary_data)
summary(regr)
```

```{r, results="hold"}
# Extract the first 5 fitted values from the model
fitted_values <- head(fitted(regr), 5)
print("First 5 Fitted Values:")
print(fitted_values)

# Extract the first 5 residuals from the model
residuals <- head(residuals(regr), 5)
print("First 5 Residuals:")
print(residuals)
```

## 2b

Use only linear algebra and the geometric view of regression to estimate the regression yourself

### (i)(ii)

```{r, results="hold"}
# Create an X matrix that has a first column of 1s followed by columns of the independent variables
X <- cbind(1, salary_data$Experience, salary_data$Score, salary_data$Degree)

# Create a y vector with the Salary values
y <- salary_data$Salary
```

### (iii)

```{r, results="hold"}
# Compute the beta_hat vector of estimated regression coefficients 
beta_hat <- solve(t(X) %*% X) %*% t(X) %*% y
beta_hat
```

### (iv)

```{r, results="hold"}
# Using the above, compute a y_hat vector of estimated y values, and a res vector of residuals 
y_hat <- X %*% beta_hat
res <- y - y_hat

# Show the code and the first 5 values of y_hat and res
print(head(y_hat, 5))
print(head(res, 5))
```

### (v)

```{r, results="hold"}
# Using only the results from (i) – (iv), compute SSR, SSE and SST
SSR <- sum((y_hat - mean(y))^2)
SSE <- sum((y - y_hat)^2)
SST <- sum((y - mean(y))^2)

cat("SSR:", SSR, "\n")
cat("SSE:", SSE, "\n")
cat("SST:", SST, "\n")
```

## 2c

Compute R\^2 for in two ways, and confirm you get the same results

### (i)

Use any combination of SSR, SSE, and SST

```{r, results="hold"}
SSR / SST
```

### (ii)

Use the squared correlation of vectors y and y_hat

```{r, results="hold"}
cor(y, y_hat)^2
```

[**Observation**]{.underline}

Two methods yield the same result indeed.

# Question 3

We're going to look back at the early, heady days of global car manufacturing, when American, Japanese, and European cars competed to rule the world. Take a look at the data set in file auto-data.txt. We are interested in explaining what kind of cars have higher fuel efficiency (mpg).

| **mpg/** miles-per-gallon (dependent variable)
| **cylinders/** cylinders in engine
| **displacement/** size of engine
| **horsepower/** power of engine
| **weight/** weight of car
| **acceleration/** acceleration ability of car
| **model_year/** year model was released
| **origin/** place car was designed (1: USA, 2: Europe, 3: Japan)
| **car_name/** make and model names
| 

```{r, results="hold"}
auto <- read.table("auto-data.txt", header=FALSE, na.strings = "?")
names(auto) <- c("mpg", "cylinders", "displacement", "horsepower", "weight", 
                 "acceleration", "model_year", "origin", "car_name")
```

## 3a

Let's first try exploring this data and problem.

### (i)

Visualize the data as you wish (report only relevant/interesting plots)

```{r, results="hold"}
library(ggplot2)

vars <- c("cylinders","displacement", "horsepower", "weight", "acceleration","model_year","origin")
for (var in vars) {
  p <- ggplot(auto, aes_string(x = var, y = "mpg")) +
    geom_point(alpha = 0.5, color = "grey",cex = 2) +
    labs(title = paste("mpg vs", var), x = var, y = "mpg")
  print(p)
}

```

### (ii)

Report a correlation table of all variables, rounding to two decimal places

```{r, results="hold"}
# Compute the correlation matrix
cor_matrix <- cor(auto[, sapply(auto, is.numeric)], use = "pairwise.complete.obs")
round(cor_matrix, 2)
```

### (iii)

[**Question**]{.underline}

From the visualizations and correlations, which variables appear to relate to mpg?

[**Answer**]{.underline}

displacement, horsepower and weight

### (iv)

[**Question**]{.underline}

Which relationships might not be linear?

[**Answer**]{.underline}

cylinders, model_year, origin

### (v)

[**Question**]{.underline}

Are there any pairs of independent variables that are highly correlated (r \> 0.7)?

[**Answer**]{.underline}

cylinders, displacement, horsepower, and weight

## 3b

Let's create a linear regression model where mpg is dependent upon all other suitable variables

### (i)

[**Question**]{.underline}

Which independent variables have a 'significant' relationship with mpg at 1% significance?

```{r, results="hold"}
# Fit the linear regression model
regr <- lm(mpg ~ cylinders + displacement + horsepower + weight + acceleration + model_year + factor(origin), data = auto)
summary(regr)
```

[**Answer**]{.underline}

weight, model_year

### (ii)

[**Question**]{.underline}

Looking at the coefficients, is it possible to determine which independent variables are the most effective at increasing mpg? If so, which ones, and if not, why not? (hint: units!)

[**Answer**]{.underline}

No. The data are comparable only when they are standardized, or we are simply juggling with different units.

## 3c

Let's try to resolve some of the issues with our regression model above.

### (i)

[**Question**]{.underline}

Create fully standardized regression results: are these slopes easier to compare?

```{r, results="hold"}
auto_std <- data.frame(cbind(scale(auto[,1:7]), origin = auto[,8]))
regr_std <- lm(mpg ~ cylinders + displacement + horsepower + weight + acceleration + model_year + factor(origin), data = auto_std)
summary(regr_std)
```

[**Answer**]{.underline}

Yes. All variables get uniform units, so they become comparable.

### (ii)

[**Question**]{.underline}

Regress mpg over each nonsignificant independent variable, individually. Which ones become significant when we regress mpg over them individually?

[**Observation**]{.underline}

According to the summary table above, three variables (cylinders, horsepower, and acceleration) are not significant as their p-value are greater than 0.01.

```{r, results="hold"}
regr_cylinders <- lm(mpg ~ cylinders, data = auto_std)
summary(regr_cylinders)
```

```{r, results="hold"}
regr_horsepower <- lm(mpg ~ horsepower, data = auto_std)
summary(regr_horsepower)
```

```{r, results="hold"}
regr_acceleration <- lm(mpg ~ acceleration, data = auto_std)
summary(regr_acceleration)
```

### (iii)

[**Question**]{.underline}

Plot the distribution of the residuals: are they normally distributed and centered around zero?\
(get the residuals of a fitted linear model, e.g. regr \<- lm(...), using regr\$residuals

```{r, }
residual1 <- data.frame(residual = regr_std$residuals)
residual2 <- data.frame(residual = regr_cylinders$residuals)
residual3 <- data.frame(residual = regr_horsepower$residuals)
residual4 <- data.frame(residual = regr_acceleration$residuals)
```

```{r, results="hold"}

plot1 <- ggplot(x = residual1)+
  geom_histogram(aes(y = after_stat(density)), color = "grey" , bins = 40)+
  labs(title = "STD Residuals Histogram", x = "residuals", y = "frequency")+
  geom_density(aes(x = residual1), color = "darkgrey", lwd = 1)
plot2 <- ggplot(x = residual2)+
  geom_histogram(aes(y = after_stat(density)), color = "grey" , bins = 40)+
  labs(title = "CYLINDERS Residuals Histogram", x = "residuals", y = "frequency")+
  geom_density(aes(x = residual2), color = "darkgrey", lwd = 1)
plot3 <- ggplot(x = residual3)+
  geom_histogram(aes(y = after_stat(density)), color = "grey" , bins = 40)+
  labs(title = "HORSEPOWER Residuals Histogram", x = "residuals", y = "frequency")+
  geom_density(aes(x = residual3), color = "darkgrey", lwd = 1)
plot4 <- ggplot(x = residual4)+
  geom_histogram(aes(y = after_stat(density)), color = "grey" , bins = 40)+
  labs(title = "ACCELERATION Residuals Histogram", x = "residuals", y = "frequency")+
  geom_density(aes(x = residual4), color = "darkgrey", lwd = 1)
```

[**Answer**]{.underline}

According to the graphs presented above, the residuals are clearly normally distributed and centered around zero as well.
